---
title: "OMOP Data Quality Checks Report on CCAA"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document:
    self_contained: true
    toc: true
    toc_depth: 2
editor_options: 
  chunk_output_type: console
params:
  driver:
    label: "Driver:"
    value: PostgreSQL
    input: select
    choices: [PostgreSQL, ODBC, SQLite]
  host: 
    label: "Host:"
    value: default_host
    input: text
  port: 
    label: "Port:"
    value: 5432
    input: numeric
  dbname: 
    label: "Database Name:"
    value: database
    input: text
  schema: 
    label: "Schema:"
    value: decovid_omop
    input: text
  user: 
    label: "User name:"
    value: user_name
    input: text
  password: 
    label: "Password:"
    value: password
    input: password
  local_hospital:
    label: "Hospital name:"
    value: CCAA
    input: text
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = FALSE, warning = FALSE, error = FALSE,
  out.width = "100%",
  dpi = 300
)
```

```{r setup}
library(knitr)
library(ggplot2)
library(dplyr)
library(gt)
library(patchwork)
library(markovchain)
library(stringr)
library(readr)
library(DBI)
```

```{r prepare-st}
ctn <- setup_ctn(params)
schema <- params$schema

st <- prepare_tables(ctn, schema)
overview <- prepare_overview(st)

vo_id_true <- all(!is.na(overview$visit_occurrence_id))

raw_sql <- read_file("measurement_observation_summary.sql") %>% 
    str_replace_all(":OMOP_SCHEMA", "cca_omop")
measurement_observation_summary <- dbGetQuery(ctn, raw_sql) %>%
    as_tibble()

raw_sql <- read_file("condition_procedure_summary.sql") %>% 
    str_replace_all(":OMOP_SCHEMA", "cca_omop")
condition_procedure_summary <- dbGetQuery(ctn, raw_sql) %>%
    as_tibble()

raw_sql <- read_file("drug_summary.sql") %>% 
    str_replace_all(":OMOP_SCHEMA", "cca_omop")
drug_summary <- dbGetQuery(ctn, raw_sql) %>%
    as_tibble()
```


Check that all visit_occurrence_id is populated: `r vo_id_true`

```{r}
clinical_tbls <- prepare_tally(ctn, schema)

zero_vals <- clinical_tbls %>%
  filter(n == 0) %>%
  select(table) %>%
  pull()
```

## Person Table

```{r ages}
if ("person" %in% zero_vals) {
  cat("There is no data in the person table")
} else {
  check_age <- min(overview$age, na.rm = TRUE) >= 18
}
```
Madatory checks

- Minimum age >= 0: `r check_age`


## Visit Characteristics

Date boundaries of the cohort:

```{r cohort-boundaries, echo=FALSE, warning=FALSE}
if ("visit_occurrence" %in% zero_vals) {
  cat("There is no data in the visit_occurrence table")
} else {
  st[["visit_occurrence"]] %>%
    prepare_cohort_boundaries() %>%
    gt() %>%
    tab_style(
      style = cell_fill(color = "red1", alpha = 0.5),
      locations = cells_body(
        rows = tolerance == "fail")
    ) %>%
    tab_options(table.width = pct(100))
}
```

All deaths inside the hospital should be aligned with the final visit end datetime.
This is an OMOP convention and doesn't necessarily mean the data is bad.

```{r ghost-visit}
ghost_visit <- st[["visit_occurrence"]] %>%
  left_join(st[["death"]] %>%
              select(person_id, death_date, death_datetime),
            by = "person_id") %>%
  filter(visit_start_date > death_date)

#print(testthat::expect_true(nrow(ghost_visit)))
print(nrow(ghost_visit))
```

All records that have a discharge recorded at "Patient died" should also have a death record in the death table.

```{r death-discharge-agreement}
st[["visit_occurrence"]] %>%
  filter(discharge_to_concept_id == "Patient died") %>%
  left_join(
    st[["death"]] %>%
      select(person_id, death_date, death_datetime) %>%
      mutate(death_record = TRUE),
    by = "person_id") %>%
  group_by(death_record = !is.na(death_record)) %>%
  tally() %>%
  gt()
```

Deviations of more than 24 hours (by datetime) or 48 hours (by date) should be investigated:

```{r death_discrep}
death_discrep <- overview %>%
  filter(
    !is.na(visit_concept_id) &
    (!is.na(death_datetime) | !is.na(death_date)) &
    discharge_to_concept_id == "Patient died") %>%
  group_by(person_id) %>%
  filter(max(visit_start_datetime) == visit_start_datetime) %>%
  ungroup() %>%
  select(person_id, death_datetime, death_date, visit_end_datetime) %>%
  mutate(
    diff_date = as.integer(difftime(visit_end_datetime, death_date, units = "hours")),
    diff_dttm = as.integer(difftime(visit_end_datetime, death_datetime, units = "hours")))

death_discrep %>%
  summarise(
    `death 24 hrs < visit end (dttm)` = sum(diff_dttm > 24, na.rm = TRUE),
    `death within tolerance (dttm)` = sum(diff_dttm <= 24 & diff_dttm >= -24, na.rm = TRUE),
    `death 24 hrs > visit end (dttm)` = sum(diff_dttm < -24, na.rm = TRUE),
    `death 48 hrs < visit end (date)` = sum(diff_date > 48, na.rm = TRUE),
    `death within tolerance (date)` = sum(diff_date <= 48 & diff_date >= -48, na.rm = TRUE),
    `death 48 hrs > visit end (date)` = sum(diff_date < -48, na.rm = TRUE),
  ) %>%
  tidyr::pivot_longer(everything(), names_to = "tolerance test", values_to = "n") %>%
  group_by(dttm = grepl(x = `tolerance test`, pattern = "dttm")) %>%
  mutate(
    total = sum(n),
    `%` = round((n/total)*100, 0)) %>%
  ungroup() %>%
  mutate(tolerance = c(5, 90, 5, 5, 90, 5)) %>%
  select(-dttm, -total) %>%
  gt() %>%
    tab_style(
      style = cell_fill(color = "red1", alpha = 0.5),
      locations = cells_body(
        rows = (`%` < tolerance & grepl("tolerance", `tolerance test`)) |
               (`%` > tolerance & !grepl("tolerance", `tolerance test`)))
    ) %>%
  tab_options(table.width = pct(100))
```


```{r death-discrepancy-plot}
plot_min <- min(pmin(death_discrep$diff_date, death_discrep$diff_dttm), na.rm = TRUE)
plot_max <- max(pmax(death_discrep$diff_date, death_discrep$diff_dttm), na.rm = TRUE)

death_discrep_date <- death_discrep %>%
  filter(!is.na(diff_date)) %>%
  ggplot(aes(x = diff_date)) +
  geom_density() +
  theme_classic() +
  ggtitle("Hours difference between visit end and death date") +
  theme(
    axis.text.y = element_blank(),
    axis.title = element_blank()) +
  lims(x = c(plot_min, plot_max))

death_discrep_dttm <- death_discrep %>%
  filter(!is.na(diff_dttm)) %>%
  ggplot(aes(x = diff_dttm)) +
  geom_density() +
  theme_classic() +
  labs(x = "time difference (hours)") +
  ggtitle("Hours difference between visit end and death date time") +
  theme(
    axis.text.y = element_blank(),
    axis.title.y = element_blank()) +
  lims(x = c(plot_min, plot_max))

death_discrep_date / death_discrep_dttm
```

## Visit Occurrence

Are all visits ordered sequentially:

```{r overlapping-vo}
overlapping <- overview %>%
  select(person_id, visit_start_datetime, visit_end_datetime) %>%
  group_by(person_id) %>%
  arrange(person_id, visit_start_datetime) %>%
  mutate(lead_start = quick_lead(visit_start_datetime)) %>%
  mutate(overlap = lead_start < visit_end_datetime) %>%
  mutate(overlap = if_else(is.na(overlap), FALSE, overlap))

overlapping %>%
  group_by(overlap) %>%
  tally() %>%
  gt()
```

What are the lengths of stay:

```{r length-of-stay-vo}
res_adm <- st[["visit_occurrence"]] %>%
  mutate(
    los = as.numeric(
        difftime(visit_end_datetime, visit_start_datetime, units = "days"))) %>%
  select(person_id, los)

los_plot <- res_adm %>%
  filter(!is.na(los)) %>%
  mutate(outlier = outliers(los)) %>%
  mutate(los = if_else(
    outlier == "main", los, scales::rescale(los, to = c(max(los)-0.02*max(los), max(los))))) %>%
  ggplot(aes(x = los)) +
  geom_density() +
  facet_grid(cols = vars(outlier), scales = "free_x", space = "free_x") +
  theme_classic() +
  theme(strip.background = element_rect(fill = "grey80", colour = "grey80")) +
  labs(x = "length of stay (days)")

los_tab <- ggplotGrob(los_plot)
p_filtered <- gtable_filter_remove(los_tab, name = "axis-b-2",
                                   trim = TRUE)

grid::grid.newpage()
grid::grid.draw(p_filtered)
```


```{r negative-los-vo}
res_adm_tbl <- res_adm %>%
  mutate(
    `length of stay` = case_when(
      los < 0 ~ "negative",
      los == 0 ~ "zero",
      los > 0 ~ "positive",
      TRUE ~ "missing los data"
    )
  ) %>%
  group_by(`length of stay`) %>%
  tally() %>%
  mutate(
    total = sum(n),
    `%` = round((n/total)*100, 0)
  ) %>%
  select(-total)

res_adm_toll <- tibble(
  `length of stay` = c("negative", "zero", "positive", "missing los data"),
  tolerance = c(1, 1, 99, 1)
)

res_adm_toll %>%
  left_join(res_adm_tbl, by = "length of stay") %>%
  gt() %>%
  tab_style(
    style = cell_fill(color = "red1", alpha = 0.5),
    locations = cells_body(
      rows = (`%` > tolerance & `length of stay` != "positive") |
             (`%` < tolerance & `length of stay` == "positive")
  )) %>%
  tab_options(table.width = pct(100))
```

## Visit Detail

Are all visits ordered sequentially:

```{r overlapping-vd}
overlapping <- st[["visit_detail"]] %>%
  select(person_id, visit_detail_start_datetime, visit_detail_end_datetime) %>%
  group_by(person_id) %>%
  arrange(person_id, visit_detail_start_datetime) %>%
  mutate(lead_start = quick_lead(visit_detail_start_datetime)) %>%
  mutate(overlap = lead_start < visit_detail_end_datetime) %>%
  mutate(overlap = if_else(is.na(overlap), FALSE, overlap))

overlapping %>%
  group_by(overlap) %>%
  tally() %>%
  gt()
```

What are the lengths of stay:

```{r length-of-stay-vd}
cs <- tbl(ctn, in_schema(schema, "care_site")) %>%
  collect() %>%
  select(
    concept_id = care_site_id,
    concept_name = care_site_name
  ) %>%
  mutate(concept_id = as.integer(concept_id))

st[["visit_detail"]]$care_site_id <- match_concepts(st[["visit_detail"]]$care_site_id, cs)

## Temporary fix

st[["visit_detail"]]$care_site_id <-
  if_else(is.na(st[["visit_detail"]]$care_site_id),
          "Out Patient Department",
          st[["visit_detail"]]$care_site_id)

res_adm <- st[["visit_detail"]] %>%
  mutate(
    los = as.numeric(
        difftime(visit_detail_end_datetime, visit_detail_start_datetime, units = "days"))) %>%
  select(person_id, care_site_id, los)

res_adm_tbl <- res_adm %>%
  mutate(
    `length of stay` = case_when(
      los < 0 ~ "negative",
      los == 0 ~ "zero",
      los > 0 ~ "positive",
      TRUE ~ "missing los data"
    )
  ) %>%
  group_by(`length of stay`) %>%
  tally() %>%
  mutate(
    total = sum(n),
    `%` = round((n/total)*100, 0)
  ) %>%
  select(-total)

res_adm_toll <- tibble(
  `length of stay` = c("negative", "zero", "positive", "missing los data"),
  tolerance = c(1, 1, 99, 1)
)

res_adm_toll %>%
  left_join(res_adm_tbl, by = "length of stay") %>%
  gt() %>%
  tab_style(
    style = cell_fill(color = "red1", alpha = 0.5),
    locations = cells_body(
      rows = (`%` > tolerance & `length of stay` != "positive") |
             (`%` < tolerance & `length of stay` == "positive")
  )) %>%
  tab_options(table.width = pct(100))
```

## Measurements

Missing concepts are highlighted in **red**:

```{r measurement1}
meas <- tbl(ctn, in_schema(schema, "measurement")) %>%
  group_by(measurement_concept_id) %>%
  tally() %>%
  collect() %>%
  mutate(across(everything(), as.integer))

# Measurements that are in the roadmap & the DB
meas_in_ref <- meas$measurement_concept_id[meas$measurement_concept_id %in% dq_ref$concept_id]

# Measurements that are not in the roadmap but appear in the DB
meas_nin_ref <- meas$measurement_concept_id[!(meas$measurement_concept_id %in% dq_ref$concept_id)]

# Measurements that appear in the DB, but are not in the roadmap
meas_out <- dq_ref$concept_id[!(dq_ref$concept_id %in% meas$measurement_concept_id)]

meas_all <- sort(unique(c(meas_in_ref, meas_nin_ref, meas_out)))
meas_all <- mini_dict(ctn, schema, meas_all)

meas_all <- meas_all %>%
  mutate(
    status = case_when(
      concept_id %in% meas_in_ref ~ "Onboarded",
      concept_id %in% meas_out ~ "Pending",
      concept_id %in% meas_nin_ref ~ "Not in roadmap"
  )) %>%
  select(concept_id, concept_name, status) %>%
  arrange(concept_id) %>%
  left_join(meas, by = c("concept_id" = "measurement_concept_id")) %>%
  mutate(n = if_else(is.na(n), 0L, n))

meas_all %>%
  gt() %>%
  tab_style(
    style = cell_fill(color = "red1", alpha = 0.5),
    locations = cells_body(
      rows = status == "Pending")
  ) %>%
  tab_options(table.width = pct(100))
```

This section details all measurements:

Error codes:

* **before**: sample taken before visit
* **after**: sample taken after visit
* **no visit**: sample taken without an associated visit
* **duplications**: suspicious for duplicate samples


```{r measurements2}
meas_dq <- meas_all %>%
  left_join(dq_ref[,c("concept_id", "short_name", "target_column",
                      "units_concept_id", "limits")],
            by = "concept_id")

rmd <- vector(mode = "list", length = nrow(meas_dq))

for (i in seq_along(meas_dq$concept_id)) {
  if (meas_dq$status[i] == "Onboarded") {
      rmd[[i]] <- try(knitr::knit_expand("child_measurement.Rmd",
                                 x = meas_dq$concept_id[i]))
  } else if (meas_dq$status[i] == "Pending") {
      rmd[[i]] <- try(knitr::knit_expand("child_pending.Rmd",
                                 x = meas_dq$concept_id[i]))
  } else {
      rmd[[i]] <- try(knitr::knit_expand("child_off_roadmap.Rmd",
                                 x = meas_dq$concept_id[i]))
  }

}

rmd <- meas_dq$concept_id %>%
  purrr::map(
    ~ knitr::knit_expand("child_measurement.Rmd", x = .x)
  )

rmd <- paste(rmd, collapse = "\n")
```

```{r results = "asis"}
rendered <- knit(text = rmd, quiet = TRUE)
cat(rendered, sep = "\n")
```



```{r disconnect}
DBI::dbDisconnect(ctn)
```











