---
title: "OMOP Summary Report on CCAA"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document:
    self_contained: true
    toc: true
    toc_depth: 2
editor_options: 
  chunk_output_type: console
params:
  driver:
    label: "Driver:"
    value: PostgreSQL
    input: select
    choices: [PostgreSQL, ODBC, SQLite]
  host: 
    label: "Host:"
    value: default_host
    input: text
  port: 
    label: "Port:"
    value: 5432
    input: numeric
  dbname: 
    label: "Database Name:"
    value: database
    input: text
  schema: 
    label: "Schema:"
    value: decovid_omop
    input: text
  user: 
    label: "User name:"
    value: user_name
    input: text
  password: 
    label: "Password:"
    value: password
    input: password
  local_hospital:
    label: "Hospital name:"
    value: CCAA
    input: text
---

```{r, include = FALSE}
library(knitr)
library(ggplot2)
library(dplyr)
library(patchwork)
library(readr)
library(DBI)

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = FALSE, warning = FALSE, error = FALSE,
  out.width = "100%",
  dpi = 300
)
```

## Structural Checks

```{r prepare-st}

# Connect to the database. 
# Read core tables and prepare overview. 
ctn <- setup_ctn(params)

st <- prepare_tables(ctn, params$schema)
overview <- prepare_overview(st)

# Getting counts of all the tables.
table_counts <- prepare_tally(ctn, params$schema)

zero_vals <- table_counts %>%
  filter(n == 0) %>%
  select(table) %>%
  pull()

measurement_observation_summary <- read_file(system.file("measurement_observation_summary.sql", 
                                                         package = "ccaa.inspectEHR")) %>% 
    str_replace_all(":OMOP_SCHEMA", params$schema) %>% 
  dbGetQuery(ctn, .) %>%
    as_tibble()

condition_procedure_summary <- read_file(system.file("condition_procedure_summary.sql", package = "ccaa.inspectEHR")) %>% 
    str_replace_all(":OMOP_SCHEMA", params$schema) %>% 
  dbGetQuery(ctn, .) %>%
    as_tibble()

drug_summary <- read_file(system.file("drug_summary.sql", package = "ccaa.inspectEHR")) %>% 
    str_replace_all(":OMOP_SCHEMA", params$schema) %>% 
  dbGetQuery(ctn, .) %>%
    as_tibble()

DBI::dbDisconnect(ctn)
```

```{r row-counts}
summary <- tibble(
  variable = character(),
  value = character())

summary <- bind_rows(summary, c(c(variable = "Number of patients"), 
                                value = filter(table_counts, table == "person") %>% pull(n)))
summary <- bind_rows(summary, c(c(variable = "Number of visit occurrences"), 
                                value = filter(table_counts, table == "visit_occurrence") %>% pull(n)))
summary <- bind_rows(summary, c(c(variable = "Number of care sites"), 
                                value = filter(table_counts, table == "care_site") %>% pull(n)))
summary <- bind_rows(summary, c(c(variable = "Number of males"), 
                                value = nrow(st[["person"]] %>% filter(gender_concept_id == "MALE"))))
summary <- bind_rows(summary, c(c(variable = "Number of females"), 
                                value = nrow(st[["person"]] %>% filter(gender_concept_id == "FEMALE"))))
summary <- bind_rows(summary, c(c(variable = "Number of missing gender records"), 
                                value = nrow(st[["person"]] %>% filter(is.na(gender_concept_id)))))
summary <- bind_rows(summary, c(c(variable = "Number of missing years of birth"), 
                                value = nrow(st[["person"]] %>% filter(is.na(year_of_birth)))))
summary <- bind_rows(summary, c(c(variable = "Number of measurements"),
                                value = filter(table_counts, table == "measurement") %>% pull(n)))
summary <- bind_rows(summary, c(c(variable = "Number of observations"), 
                                value = filter(table_counts, table == "observation") %>% pull(n)))
summary <- bind_rows(summary, c(c(variable = "Number of conditions"), 
                                value = filter(table_counts, table == "condition_occurrence") %>% pull(n)))
summary <- bind_rows(summary, c(c(variable = "Number of procedures"), 
                                value = filter(table_counts, table == "procedure_occurrence") %>% pull(n)))
summary <- bind_rows(summary, c(c(variable = "Number of drug exposures"), 
                                value = filter(table_counts, table == "drug_exposure") %>% pull(n)))
```

### General counts
```{r}
summary %>%
  mutate(value = as.integer(value)) %>% 
  print_large_kable()
```


## Person Table

```{r ages}
if ("person" %in% zero_vals) {
  cat("There is no data in the person table")
} else {
  age_plot <- plot_age(overview)
  check_age <- min(overview$age, na.rm = TRUE) >= 18

  sex_plot <- plot_sex(st[["person"]])
  ethnic_plot <- plot_ethnicity(st[["person"]])

  age_plot + sex_plot + ethnic_plot
}
```

## Visit Characteristics

Number of visits stratified by visit type:

```{r visit-types}
if (!("visit_occurrence" %in% zero_vals)) {
  st[["visit_occurrence"]] %>%
    group_by(visit_concept_id) %>%
    summarise(
      count = n(),
      percentage = round((n()/nrow(st[["visit_occurrence"]]))*100, 0)
    ) %>% 
    print_large_kable()
}
```

```{r visit-profile}
plot_visit_profile(st[["visit_occurrence"]])
```

```{r admission-from1}
asci <- vo_ans %>%
  filter(target_columnn == "admitted_from_concept_id") %>%
  select(concept_id, concept_name) %>%
  distinct()

st[["visit_occurrence"]] %>%
  filter(!(admitted_from_concept_id %in% asci$concept_name)) %>%
  group_by(admitted_from_concept_id) %>%
  tally() %>%
  print_large_kable()
```

```{r admission-from2}
st[["visit_occurrence"]] %>%
  mutate(admitted_from_concept_id = forcats::fct_infreq(as.character(admitted_from_concept_id))) %>%
  group_by(admitted_from_concept_id) %>%
  tally() %>%
  check_zero_tally(n) %>%
  ggplot(aes(x = admitted_from_concept_id)) +
  geom_point(aes(y = n, colour = .is_zero)) +
  geom_segment(aes(
    y = 0,
    yend = n,
    xend = admitted_from_concept_id)) +
  scale_colour_manual(values = c("red", "black"), drop = FALSE) +
  theme_classic() +
  labs(y = "count") +
  theme(
    legend.position = "none",
    plot.title.position = "plot",
    axis.title.y = element_blank()) +
    coord_flip() +
  ggtitle("Admission Source")
```

Discharge codes outside the approved list:

```{r discharge-to1}
dtci <- vo_ans %>%
  filter(target_columnn == "discharged_to_concept_id") %>%
  select(concept_id, concept_name) %>%
  distinct()

st[["visit_occurrence"]] %>%
  filter(!(discharged_to_concept_id %in% dtci$concept_name)) %>%
  group_by(discharged_to_concept_id) %>%
  tally() %>%
  print_large_kable()
```

```{r discharge-to2}
st[["visit_occurrence"]] %>%
  mutate(discharged_to_concept_id = forcats::fct_infreq(as.character(discharged_to_concept_id))) %>%
  group_by(discharged_to_concept_id) %>%
  tally() %>%
  check_zero_tally(n) %>%
  ggplot(aes(x = discharged_to_concept_id)) +
  geom_point(aes(y = n, colour = .is_zero)) +
  geom_segment(aes(
    y = 0,
    yend = n,
    xend = discharged_to_concept_id)) +
  scale_colour_manual(values = c("red", "black"), drop = FALSE) +
  theme_classic() +
  labs(y = "count") +
  theme(
    legend.position = "none",
    plot.title.position = "plot",
    axis.title.y = element_blank()) +
  coord_flip() +
  ggtitle("Discharge To")
```

## Outcomes

The following list the outcomes for patients attending at **`r params$local_hospital`**:

```{r outcomes}
overview %>%
  distinct(person_id, .keep_all = TRUE) %>%
  filter(!is.na(visit_occurrence_id)) %>%
  group_by(visit_concept_id, death = !is.na(death_datetime)) %>%
  tally() %>%
  group_by(visit_concept_id) %>%
  mutate(
    total = sum(n),
    `%` = round((n/total)*100, 0)
  ) %>%
  print_large_kable()
```

## Distributions of Times

Typically patients are admitted earlier in the day, and discarged later.
Death often has a uniform distribution, with perhaps a slight peak in the early morning or later afternoon.

```{r outcome-dist}
overview %>%
  select(visit_start_datetime, visit_end_datetime, death_datetime) %>%
  mutate_all(~ as.POSIXct(., tz = "UTC")) %>%
  filter_all(any_vars(!is.na(.))) %>%
  mutate_all(~ hms::as_hms(.)) %>%
  tidyr::pivot_longer(everything(), names_to = "time_point", values_to = "time") %>%
  ggplot(aes(x = time, fill = time_point)) +
  geom_density(alpha = 0.5) +
  theme_classic() +
  labs(x = "event time distribution", fill = "event") +
  theme(legend.position = "bottom")
```

### Measurements and observations
The 50 most common concepts are displayed.
```{r}
measurement_observation_summary %>%
  print_large_kable(max_rows = 50)
```

### Condition and procedure occurence.
The 50 most common concepts are displayed.
```{r}
condition_procedure_summary %>%
  print_large_kable(max_rows = 50)
```

### Drug exposure
The 50 most common concepts are displayed.
```{r}
drug_summary %>%
  print_large_kable(max_rows = 50)
```